
---
title: "Data Science Midterm Assessment"
author: "Put Your Name Here"
date: "Due: October 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(stringr)
library(dplyr)
library(ggplot2)
```

## Instructions

This is an individual assignment.  You may discuss how functions work with your classmates, but you cannot ask someone to debug/fix your code.  You have to do that!  Any failure to do so will lead to a 0 on this assignment and an Academic Honesty Violation.

Make all edits in this R Markdown Document.  No code should appear in the knitted document.  You should have one code block for each question.  Upload this .Rmd (which should knit) to eLearn by the end of Tuesday, October 17.  (I recommend working on tidying the data only when a question requires it.  To make sure you can get back to working code.  Use git and commit after you have answered a question.  Questions probably shouldn't be answered in the order they are written.)

## The Data

The schedule for next semester can be found at <http://www2.loras.edu/admin/sprsched2.html>. (Old versions can be found at <http://www2.loras.edu/admin/fallsched.html> and <http://www2.loras.edu/admin/sprsched.html>.  These maybe useful for testing purposes for some things.)

## The Tasks
1. Import the data into R and tidy it up.  (I recommend `read_fwf()` for import.  `lead()` and `lag()` will be your friends for tidying.  The [chapter on strings](http://r4ds.had.co.nz/strings.html) may also help.)

```{r}
Spring_Schedule <- read_fwf(file = "http://www2.loras.edu/admin/sprsched2.html", fwf_widths(c(3,5,6,7,5,4,5,36,10,15,8,6,6,30), c("MAX", "USED", "WTLST", "SYN", "DEPT", "CAT", "SEC", "TITLE", "CRED", "TIME", "DAY", "BLDG", "ROOM", "INSTRUCTOR")), skip = 5)
```


2. Create visualizations that summarize the following:

    a. Distribution of number of courses offered in each discipline.
    
    ```{r}
    Spring_Schedule %>%
      
        group_by(DEPT) %>%
      
            filter(!is.na(DEPT)) %>%
      
                summarise(Num_Courses_Per_Discipline = n()) %>%
      
                    ggplot() + geom_col(mapping = aes(x = DEPT, y = Num_Courses_Per_Discipline)) + coord_flip()
    ```

    b. Distribution of level (100, 200, 300, 400, ...) of courses.
    
    ```{r}
Spring_Schedule %>%
  
    group_by(CAT) %>%
  
      filter(!is.na(CAT)) %>%
  
          summarise(Category_Count = n()) %>%
  
                  ggplot() + geom_col(mapping = aes(x = CAT, y = Category_Count)) + coord_flip()
``` 

    c. Distribution of level of course by discipline.
    
    ```{r}
   Spring_Schedule %>%
      
        group_by(DEPT, CAT) %>%
      
            filter(!is.na(DEPT) & !is.na(CAT)) %>%
      
                summarise(Course_Level_Dept_Dist = n()) %>%
      
                    ggplot() + geom_col(mapping = aes(x = DEPT, y = Course_Level_Dept_Dist)) + coord_flip()
    ```

    d. Distribution of start time of courses.
    
    ```{r}
    
    Spring_Schedule %>%
      
        separate(TIME, into = c("Start_Time", "End_Time"), sep = "-") %>%
      
            group_by(Start_Time) %>%
    
               filter(!is.na(Start_Time)) %>%
      
                filter(Start_Time != "Their Own Land" & Start_Time != "t. of Latin Ame" & Start_Time != "ronmental Bio" & Start_Time != "ports" & Start_Time != "ort" & Start_Time != "nmental Politic" & Start_Time != "minar" & Start_Time != "ir. Sustain." & Start_Time != "Greece & Rome" & Start_Time != "Film: Africa" & Start_Time != "e following date" & Start_Time != "dentity Am West" & Start_Time != "ature/Ireland" & Start_Time != "A" & Start_Time != "a" & Start_Time != "ental Politic" & Start_Time != ". Sustain." & Start_Time != "ure/Ireland" & Start_Time != "t" & Start_Time != "rts" & Start_Time != "reece & Rome" & Start_Time != "on:" & Start_Time != "of Latin Ame" & Start_Time != "ntity Am West" & Start_Time != "nmental Bio" & Start_Time != "nar" & Start_Time != "lm: Africa" & Start_Time != "heir Own Land" & Start_Time != "following dates:" & Start_Time != "CA"& Start_Time != "following dates") %>%
        
                  summarise(Start_Time_Distribution = n()) %>%
    
                      ggplot() + geom_col(mapping = aes(x = Start_Time, y = Start_Time_Distribution)) + coord_flip()
      
                
         
      
    ```

    e. Distribution of the duration of class meeting.
    
    ```{r}
    Spring_Schedule %>%
      
        separate(TIME, into = c("Start_Time", "End_Time"), sep = "-")
      
           # Duration = mutate("End_Time" - "Start_Time") Need to figure out how to convert chr time to numeric time
    
           # summarise()

    ```

    f. Distribution of start time of courses by discipline. (Only show the interesting ones.) I'm classifying the interesting ones as the computer science, math, and education disciplines since they are all parts of my majors and minors. 
    
    ```{r}
    Spring_Schedule %>%
      
        group_by(TIME, DEPT) %>%
      
            separate(TIME, into = c("Start_Time", "End_Time"), sep = "-") %>%
          
                filter(!is.na(DEPT) & !is.na(Start_Time)) %>%
      
                  filter(DEPT == "L.MAT" | DEPT == "L.EDU" | DEPT == "L.CIT" & Start_Time != "A") %>%
      
                    ggplot() + geom_jitter(mapping = aes(x = DEPT, y = Start_Time))
      
           

    ```

    g. Distribution of the duration of class meeting by discipline.
    
    
    ```{r}
    Spring_Schedule %>%
    
        group_by(TIME, DEPT) %>%
      
            filter(!is.na(DEPT)) %>%
    
                separate(TIME, into = c("Start_Time", "End_Time"), sep = "-")
    
    
            #Then convert Start_Time and End_Time to numerics instead of strings somehow and find the duration by doing, End_Time - Start_Time and plotting distribution of duration (y-axis) and DEPT (x-axis). 
    
                
    ```

    h. Distribution of building usage.
    
```{r}
Spring_Schedule %>%
  
    group_by(BLDG) %>%
  
      filter(!is.na(BLDG)) %>%
  
          summarise(Building_Usage = n()) %>%
  
                  ggplot() + geom_col(mapping = aes(x = BLDG, y = Building_Usage))
  
              

``` 
    
  
    i. Distribution of course fees. (What is the course fee if nothing is listed?)
    
    I would have filtered and plotted these course fees with discipline and fee total if the TITLE had all 4 lines of information in one data cell.
    
    ```{r}
    
    Spring_Schedule %>%
      
        group_by(TITLE) %>%
      
            filter(str_detect(TITLE, "COURSE FEE:")) %>%
      
                separate(TITLE, into = c("COURSE_FEE", "FEE_TOTAL"), sep = ":") %>%
      
                    ggplot() + geom_jitter(mapping = aes(x = COURSE_FEE, y = FEE_TOTAL))
    
    ```
    
  
    j. Distribution of course fees by student.
    
    There would be points plotted but the TITLE being displayed over 4 rows for me is an issue that interferes with making a proper plot. However, this is how it would be done if the information in TITLE was all in one data cell.
    
    ```{r}
    Spring_Schedule %>%
      
        group_by(TITLE, USED) %>%
      
            filter(str_detect(TITLE, "COURSE FEE")) %>%
    
              separate(TITLE, into = c("COURSE_FEE", "FEE_TOTAL"), sep = ":") %>%
        
                  ggplot() + geom_point(mapping = aes(x = FEE_TOTAL, y = USED)) + coord_flip()
    
    
    
    ```
    
    
  
3. Create blocks that show:

    a. All the courses that have space in them. (`MAX` > `USED`)
    
    ```{r}
    Spring_Schedule %>%
      
        filter(SYN, MAX > USED)
    ```
    
    b. All the courses that are `NOT OPEN TO FIRST YEAR STUDENTS`.
    
    Based off of the table below, there are 71 classes that are not open to freshman based off of what is stated in the TITLE column. Also all of the other cells are NA because of my TITLE consolidation of rows conondrum that I talk about below in e. and i. in further detail.
    
    ```{r}
    Spring_Schedule %>%
      
        group_by(TITLE) %>%
      
             filter(str_detect(TITLE, "NOT OPEN TO FIRST YEAR STUDENTS") | str_detect(TITLE, "SOPHOMORES only") | str_detect(TITLE, "JUNIORS Only") | str_detect(TITLE, "SENIORS ONLY") | str_detect(TITLE, "JUNIORS & SENIORS ONLY"))
      
       
    ```
    
    
  
    c. All the courses that are not `NOT OPEN TO FIRST YEAR STUDENTS`.
    
    ```{r}
    Spring_Schedule %>%
      
        group_by(TITLE) %>%
      
             filter(!str_detect(TITLE, "NOT OPEN TO FIRST YEAR STUDENTS") & !str_detect(TITLE, "SOPHOMORES only") & !str_detect(TITLE, "JUNIORS Only") & !str_detect(TITLE, "SENIORS ONLY") & !str_detect(TITLE, "JUNIORS & SENIORS ONLY"))
      
       
    ```
    
 
    
    
  
    d. All the courses that have space and are not `NOT OPEN TO FIRST YEAR STUDENTS`.
    
    This is not exactly correct, but would be if "NOT OPEN TO FIRST YEAR STUDENTS" was on the same line as the rest of the information in the TITLE column.
    
    ```{r}
    Spring_Schedule %>%
      
        group_by(TITLE, MAX, USED) %>%
      
            filter((MAX > USED) & !str_detect(TITLE, "NOT OPEN TO FIRST YEAR STUDENTS"))
    ```
  
    e. All the courses that have a course fee.
        Everything is NA besides the "TITLE" becuase the "COURSE FEE" portions of the TITLE are not in the same row with the rest of the information for the class. It puts each line of the title on a new line if there is too much information in subsequent lines and I'm not sure how to consolidate all of the lines unfortunately. This is the same reason this happens in other sections of my code and tables too. The information in title is split amongst 4 lines. This does display the total number of classes with course fees though
    
    ```{r}
    Spring_Schedule %>%
      
        group_by(TITLE) %>%
        
                  filter(str_detect(TITLE, "COURSE FEE"))
    
    ```
    
  
    f. All the courses that start at a given time. (Make it clear how to specify it.)
    ```{r}
    Spring_Schedule %>%
      
      separate(TIME, into = c("Start_Time", "End_Time"), sep = "-") %>%
      
        group_by(Start_Time, DAY) %>%
            
            filter(!is.na(Start_Time) & !is.na(DAY))  %>%
    
               filter(Start_Time == "10:00")
    ```
  
    g. All the courses that are on a given day. (Make it clear how to specify it.) ---> All the given courses on Monday
    
    ```{r}
    Spring_Schedule %>%
      
        group_by(TITLE, DAY) %>%
            
            filter(!is.na(DAY) & !is.na(TITLE))  %>%
    
               filter(str_detect(DAY, "M"))
    ```
    
  
    h. All the courses that start at a given time on a given day. (Given time = 8:00 AM, and courses that have a meeting on day = Thursday)
    
    ```{r}
    Spring_Schedule %>%
      
        separate(TIME, into = c("Start_Time", "End_Time"), sep = "-") %>%
      
            group_by(Start_Time, DAY) %>%
    
               filter(!is.na(Start_Time) & !is.na(DAY)) %>%
        
                  filter(Start_Time == "08:00" & str_detect(DAY, "TH"))
    
        
      
    ```
    
  
    i. All the graduate classes.
    Everything is NA becuase the "GRADUATE STUDENTS ONLY" portions of the TITLE are not in the same row with the rest of the information for the class. It puts each line of the title if there is too much information in subsequent lines and I'm not sure how to consolidate all of the lines unfortunately. This is the same reason this happens in other sections of my code and tables too. The information in title is split amongst 4 lines. There are only 10 graduate level classes.
    ```{r}
    
    
    Spring_Schedule %>%
      
            filter(str_detect(TITLE, "GRADUATE"))
      
    ```
    
    
    
    j. All the courses with a waitlist. 
    
    I tested it on the fall schedule from last year and it worked. It only showed the waitlisted classes for that semester. If there were any waitlisted classes for this coming spring at this point in time, it would display them.
    
    ```{r}
    
    
    Spring_Schedule %>%
    
            filter(str_detect(WTLST, "Y"))
      
    ```
  
4. Do something else cool.

What I did that I thought was cool is plot the classes that each math professor will be teaching in the spring with geom_point. It shows the course levels that each math department professor is teaching in the spring.



```{r}
Spring_Schedule %>%
  
  group_by(DEPT, INSTRUCTOR, CAT) %>%
  
      filter(!is.na(DEPT) & !is.na(INSTRUCTOR) & !is.na(CAT)) %>%
  
          filter(DEPT == "L.MAT") %>%
  
              ggplot() + geom_point(mapping = aes(x = INSTRUCTOR, y = CAT))


```

